Energy Dependence of Synchrotron X-Ray Rims in Tycho's Supernova Remnant

### 2014-10-28 manuscript/paper-tycho.tex
Fits w/smaller damping lengths
New Tycho FWHM tables finished; fits with ab=0.003, 0.002;
regenerate paper tables/plots that show damping fit behavior

% Magnetic damping?
Magnetic damping immediately behind the shock offers a tantalizing alternative
explanation for filament rims that requires less extreme magnetic field
amplification, and the possibility of damping in SNR shocks has not been fully
tested \citep{pohl2005, marcowith2010}.  \citet{cassam-chenai2007} examined
radio and X-ray synchrotron rim profiles, but found that neither damping nor
synchrotron losses matched radio profiles particularly well.  \citet{araya2010}
measured rims at multiple energy bands in Cas A, but suggested that the rims
only narrow more weakly than expected and did not consider magnetic damping.
\citet{rettig2012} gave model predictions for several historical SNRs and
proposed discriminating based on filament spectra -- the expectation is that
damped spectra are softer, loss-limited harder.  Recently, \citet{ressler2014}
further extended the idea of discriminating between damped or loss-limited
rims by measuring rim width energy dependence in X-ray energies and found that
rims consistently narrowed in the remnant of SN 1006, favoring synchrotron
loss-limited rims.

% Why Tycho's SNR
To further test these models, we follow \citet{ressler2014} (hereafter,
\citetalias{ressler2014}) by measuring rim widths at multiple energies in the
remnant of Tycho's 1572 supernova (hereafter, Tycho).  Tycho exhibits an
extensive shell of synchrotron-dominated thin rims around its periphery; the
rims show very little thermal emission, consistent with expansion into a low
density ISM \citet{williams2013}.  Tycho, being a nearby galactic SNR, is close
enough that its rims are easily resolved by \textit{Chandra} spatial
resolution. A deep $750 \unit{ks}$ exposure of the entire remnant from 2009
(PI: J. Hughes) also allows us to sample the remnant rims quite finely.
Furthermore, Tycho has been a prime target of study for determining whether
hadrons are being accelerated in historic SNRs \citep[and references
therein]{morlino2012}; understanding the acceleration process will also bear
upon our understanding of CR acceleration in Tycho.

@@

% Twiddle-ables in our model (the most useful part, since just modeling
% alone don't do much)
Our model fitting explores the space of varying diffusion (a bulk
representative of the underlying MHD turbulence and what have you)
and considers the two scenarios of constant magnetic field (resulting in
so-called loss-limited rims) and a damped field that falls off exponentially
behind the forward shock, as suggested by \citet{pohl2005}.
% this might go at the front before we launch into model details.

@@

% Transport fundamentals
The energy and space distribution of electrons at a supernova remnant's forward
shock controls the synchrotron rims we see in X-ray and radio.  We assume that
diffusive shock acceleration creates a power-law distribution of electrons with
an exponential cut-off at the forward shock, located at $x = 0$ with $x > 0$
increasing downstream of the shock, given as
$f(E,x=0) = K_0 E^{-s} e^{-E/\Ecut}$.  Here $K_0$ is an arbitrary
normalization, $E$ is electron energy, and the cut-off energy $\Ecut$ is given
below in Section~\ref{sec:diffcoeff}; the spectral index $s$ is taken from
the radio index $\alpha$ as $s = 2\alpha + 1$.  \citet{zirakashvili2007} have
derived from DSA a slightly different energy spectrum with super-exponential
cut-off $e^{-(p/p_{\mt{cut}})^2}$, which may impact our results; we use a
simple exponential cut-off for simplicity and consistency with
\citetalias{ressler2014}.

@@

% Can damping explain stuff?
The next question is whether damping may also produce this effect, independent
of assumptions on the shock acceleration process.  Conceptually, if damping
controls rim widths past the cut-off energy, $|\mE|$ may slightly decrease or
even increase as rims transition from being damping controlled to being
diffusion loss-limited.  If damping is only important well below the cut-off
energy, then we may see an increase in $|\mE|$ as advection becomes important.
Although magnetic damping is able to fit observed rim widths, our model assumes
diffusive shock acceleration in computing $\Ecut$ and hence we may not use our
model fits to assess whether magnetic damping can independently account for
decreasing $|\mE|$.

@@

>>>>

% Magnetic damping?
Magnetic damping immediately behind the shock offers a tantalizing alternative
explanation for filament rims that requires less extreme magnetic field
amplification, and the possibility of damping in SNR shocks has not been fully
tested \citep{pohl2005, marcowith2010}.  \citet{cassam-chenai2007} examined
radio and X-ray synchrotron rim profiles, but found that neither damping nor
synchrotron losses matched radio profiles particularly well.  \citet{araya2010}
measured rims at multiple energy bands in Cas A, but suggested that the rims
only narrow more weakly than expected and did not consider magnetic damping.
\citet{rettig2012} gave model predictions for several historical SNRs and
proposed discriminating based on filament spectra -- the expectation is that
damped spectra are softer, loss-limited harder.  Recently, \citet{ressler2014}
further extended the idea of discriminating between damped or loss-limited
rims by measuring rim width energy dependence in X-ray energies and found that
rims consistently narrowed in the remnant of SN 1006.

% Why Tycho's SNR
To further test these models, we follow \citet{ressler2014} (hereafter,
\citetalias{ressler2014}) by measuring rim widths at multiple energies in the
remnant of Tycho's 1572 supernova (hereafter, Tycho).  Tycho exhibits an
extensive shell of synchrotron-dominated thin rims around its periphery; the
rims show very little thermal emission, consistent with expansion into a low
density ISM \citep{williams2013}.  Tycho, being a nearby galactic SNR, is close
enough that its rims are easily resolved by \textit{Chandra} spatial
resolution. A deep $750 \unit{ks}$ exposure of the entire remnant from 2009
(PI: J. Hughes) also allows us to sample the remnant rims quite finely.
Furthermore, Tycho has been a prime target of study for determining whether
hadrons are being accelerated in historic SNRs \citep[and references
therein]{morlino2012}; understanding the acceleration process will also bear
upon our understanding of CR acceleration in Tycho.

@@

% Twiddle-ables in our model (the most useful part, since just modeling
% alone don't do much)
%Our model fitting explores the space of varying diffusion (a bulk
%representative of the underlying MHD turbulence and what have you)
%and considers the two scenarios of constant magnetic field (resulting in
%so-called loss-limited rims) and a damped field that falls off exponentially
%behind the forward shock, as suggested by \citet{pohl2005}.
% this might go at the front before we launch into model details.

@@

% Transport fundamentals
The energy and space distribution of electrons at a supernova remnant's forward
shock controls the synchrotron rims we see in X-ray and radio.  We assume that
diffusive shock acceleration creates a power-law distribution of electrons with
an exponential cut-off at the forward shock, located at $x = 0$ with $x > 0$
increasing downstream of the shock, given as
\begin{equation}
    f(E, x=0) = K_0 E^{-s} e^{-E/\Ecut} .
\end{equation}
Here $K_0$ is an arbitrary normalization, $E$ is electron energy, and the
cut-off energy $\Ecut$ is given below in Section~\ref{sec:diffcoeff}; the
spectral index $s$ is taken from the radio index $\alpha$ as $s = 2\alpha + 1$.
\citet{zirakashvili2007} have derived from DSA a slightly different energy
spectrum with super-exponential cut-off $e^{-(p/p_{\mt{cut}})^2}$, which may
impact our results; we use a simple exponential cut-off for simplicity and
consistency with \citetalias{ressler2014}.

@@

% Can damping explain stuff?
The next question is whether damping may also produce this effect, independent
of assumptions on the shock acceleration process.  Conceptually, if damping
controls rim widths past the cut-off energy, $|\mE|$ may slightly decrease or
even increase as rims transition from being damping controlled to being
diffusion loss-limited.  If damping is only important well below the cut-off
energy, then we may see an increase in $|\mE|$ as advection becomes important.
Although magnetic damping is able to fit observed rim widths, our model assumes
diffusive shock acceleration in computing $\Ecut$ and hence we may not use our
model fits to assess whether magnetic damping can independently account for
decreasing $|\mE|$.  This applies for loss-limited fits as well; since $\eta_2$
varies freely, $\Ecut$ may be anywhere -- and thus the model behavior is not
necessarily physically reasonable.

@@

\begin{figure}
    \centering
    \iftoggle{manuscript}{
        \includegraphics[width=0.48\textwidth]{figures/full-std_err-mE-damp-all.pdf}
        \includegraphics[width=0.48\textwidth]{figures/full-std_err-mE-damp-log.pdf}
    }{
        \plotone{figures/full-std_err-mE-damp-all.pdf}
        \plotone{figures/full-std_err-mE-damp-log.pdf}
    }
    \caption{Curves of $\mE$ values for best damping fits for all regions.
        Note the sharp contrast between most curves, and a small minority that
        blow up at low energies.
    \label{fig:mE}}
\end{figure}
